Describing demographic models is hard

### 2020-05-20 paper.tex
Update readme with GitHub repos.

\title{Describing demographic models is hard}
\author[ ]{A permutation of}
\author[1]{Aaron P. Ragsdale}
\author[1]{Dominic Nelson}
\author[2]{Jerome Kelleher}
\author[1]{Simon Gravel}
\affil[1]{McGill University and Genome Qu\'{e}bec Innovation Centre,
McGill University, Montr\'{e}al, Qu\'{e}bec, Canada}
\affil[2]{Big Data Institute, Li Ka Shing Centre for Health Information and Discovery,
University of Oxford, Oxford, United Kingdom}
\maketitle

@@

The \msprime\ coalescent simulator~\citep{kelleher2016efficient,kelleher2020coalescent}
is now quite widely used. The large increases in efficiency over the classical
\ms\ program~\citep{hudson2002generating} make it feasible to simulate large
samples of whole chromosomes for the first time. Another distinct advantage
of \msprime\ is the Python API that is its primary interface, greatly
increasing the flexibility and ease of use over the standard approach of
text-based command line interfaces. In particular, programs like \ms\
required users to specify cryptic command line options to describe demographic
models \jkcomment{maybe include an example ms command line?}. Particularly for
the large models we use today, these are not intuitive or comprehensible for humans.
The Python API for \msprime\ is a great improvement, allowing the user to
state models in a documented and reproducible manner. \jkcomment{maybe show
the same model as described above in msprime notation?}

@@

To illustrate the demography API, \msprime\ included a description of a widely-used
three population Out-of-Africa model~\citep{gutenkunst2009inferring}
as part of its tutorial documentation. In this model (Fig.~\ref{fig:ooa_stats}A),
Eurasian (CEU and CHB) and African (YRI) populations split from each other in the deep past,
followed by a more recent split of European and Asian populations, with variable rates of
continuous migration between each of the populations. However, the implementation in the
\msprime\ tutorial was incorrect. Namely, before the time of the split of African and Eurasian
populations, when there should have been just a single randomly mating population, migration was
allowed to occur between the ancestral population and a second population with size equal to
the Eurasian bottleneck size for all time into the past (Fig.~\ref{fig:ooa_stats}B).

%migration between those two populations was allowed to continue for all time into the
%past. Even though the population split was implemented correctly, the specified demographic model had 
%an additional population prior to the earliest split exchanging migrants with the ancestral population 
%(Fig.~\ref{fig:ooa_stats}B) but not contributing otherwise to present-day populations.

>>>>

\title{Describing demographic models is hard}
\author[1]{Aaron P. Ragsdale}
\author[1]{Dominic Nelson}
\author[1,$\star$]{Simon Gravel}
\author[2,$\star$]{Jerome Kelleher}
\affil[1]{McGill University and Genome Qu\'{e}bec Innovation Centre,
McGill University, Montr\'{e}al, Qu\'{e}bec, Canada}
\affil[2]{Big Data Institute, Li Ka Shing Centre for Health Information and Discovery,
University of Oxford, Oxford, United Kingdom}
\affil[$\star$]{Joint senior authors, listed alphabetically}
\maketitle

@@

The \msprime\ coalescent
simulator~\citep{kelleher2016efficient,nelson2020accounting,kelleher2020coalescent}
is now widely used. The large increases in efficiency over the classical
\ms\ program~\citep{hudson2002generating} make it feasible to simulate large
samples of whole chromosomes for the first time. Another distinct advantage
of \msprime\ is the Python API that is its primary interface, greatly
increasing the flexibility and ease of use over the standard approach of
text-based command line interfaces. In particular, programs like \ms\
required users to specify cryptic command line options to describe demographic
models \jkcomment{maybe include an example ms command line?}. Particularly for
the large models we use today, these are not intuitive or comprehensible for humans.
The Python API for \msprime\ is a great improvement, allowing the user to
state models in a documented and reproducible manner. \jkcomment{maybe show
the same model as described above in msprime notation?}

@@

To illustrate the demography API, \msprime\ included a description of a widely-used
three population Out-of-Africa model~\citep{gutenkunst2009inferring}
as part of its tutorial documentation. In this model (Fig.~\ref{fig:ooa_stats}A),
Eurasian (CEU and CHB) and African (YRI) populations split from each other in the deep past,
followed by a more recent split of European and Asian populations, with variable rates of
continuous migration between each of the populations. However, the implementation in the
\msprime\ tutorial was incorrect. Before the time of the split of African and Eurasian
populations, when there should have been just a single randomly mating population, migration was
allowed to occur between the ancestral population and a second population with size equal to
the Eurasian bottleneck size for all time into the past (Fig.~\ref{fig:ooa_stats}B).

It was used in a number of publications
~\citep{martin2017human,albers2020dating,kelleher2019inferring,tong2020population,hubisz2019mapping,
overcast2019integrated,zhou2018popdemog,pfaffelhuber2020choose}
